---
title: "Plot Simplices"
author: "Fred Kaesmann"
date: "5/20/2020"
output: pdf_document
---

```{r include=FALSE}
knitr::opts_chunk$set(
  message = FALSE,
  warning = FALSE,
  fig.width = 6,
  fig.asp = 0.618
)
```


```{r packages}
##### Packages ####
library(FredsVietorisRips)
```


### Simplices, 0 and 1 

To create our 1-simplices,
\begin{enumerate}
  \item[Step 1:] We generate a data frame of random $x$ and $y$ 
coordinates 

  \item[Step 2:] We create a matrix of pairwise distances. 
                  Each cell $i,j$ represents the distance from the $i$th point
                  to the $j$th point.
  
  \item[Step 3:] Given a distance, $\epsilon,$ we return an 
                  adjacency matrix which displays a 1 for each pair whose 
                  Euclidean Distance is within the given $\epsilon$ value
  
  \item[Step 4:] From the adjacency matrix, we pull out the points which are 
                  adjacent.
                  This is a list of all 1-simplexes in the dataset
                  
  \item[Step 5:] Last, we visualize the simplices using the `ggplot2` library
\end{enumerate}


```{r step_1}
##### Step 1: Create Data Frame ####
frame_size <- 100
df <- data.frame(
  x = runif(frame_size),
  y = runif(frame_size),
  Point = paste0("P", c(1:frame_size))
)

##### Step 2: Pairwise Distance Matrix ####

pwdmat <- Pairwisedist(df$x, df$y)
# pwdmat

##### Step 3: Given epsilon, determine Adjacency ####
# Given epsilon
epsilon <- 0.1

# Determine Adjacency 
adjacency_matrix <- AdjacencyMatrix(pwdmat, epsilon)
# adjacency_matrix


##### Step 4: Which Points are Adjacent? ####
paired_points <- AdjacentPairs(adjacency_matrix)
# paired_points

##### Step 5: Plot 0 and 1 Simplices ####
Plot_1_Simplices(df$x, df$y, epsilon)


```
