---
title: "DB Scan Pseudo Code"
output: html_notebook
---

```{r packages}
library(FredsVietorisRips)
library(magrittr)
```

Given a data set, a maximum distance for adjacency $\epsilon$, and a minimum
number of connections for a point to not be considered noise, we try to count
clusters.
```{r data}
x <- c(runif(10), runif(10) + 10, runif(10) + 20)
y <- c(runif(10), runif(10) + 10, runif(10) + 20)
```

```{r epsilon}
epsilon <- 0.5
```

We want an adjacency matrix from our given epsilon to make it easier to sort
through everything.
```{r construct_adjacency_matrix}
adjacency_matrix <- Pairwisedist(x, y) %>%
  AdjacencyMatrix(epsilon)
```


We want each point in our cluster to have a minimum number of connections to 
other points, otherwise we classify it as noise.

```{r minimum_connection_number}
min_connections <- 3
```

```{r}
l <- list()
adjmat_size <- ncol(adjacency_matrix)
v <- vector()
for (i in c(1:adjmat_size)) {
  if (sum(adjacency_matrix[i, ]) >= min_connections){
    for (j in c(1:adjmat_size)) {
      if (adjacency_matrix[i, j] == 1) {
        v <- append(v, j)
      }
    }
    for (element in v){
      if (v)
    }
  }
}
```

```{r}
##### Initialize Container ####
lmajor <- list()

adjmat_size <- nrow(adjacency_matrix)
##### Iterate ####
  for (i in c(1:adjmat_size)) {
    v <- vector()
    if (sum(adjacency_matrix[i, ]) >= min_connections) { # if it clusters
      
      for (j in c(1:length(adjacency_matrix))) {
        if (adjacency_matrix[i, j] == 1) {
          
          if (sum(adjacency_matrix[, j]) >= min_connections) {
            v <- append(v, j)
          }
        }
      }
      
      
    }
    # insert if here to check if i already in list

    if (length(v) > 0) {
      lmajor[[i]] <- v
    }
    
  } #endfori
# }
lmajor
```


```{r}
component <- list()
for (i in c(1:length(lmajor))) {
  ##### Populate list with possible clusters ####
  if (length(lmajor[[i]]) >= min_connections) {
    component[[i]] <- lmajor[[i]]
  }#endif
}#endfor
component <- Filter(Negate(is.null), component)
component
```

```{r}
for (i in c(1:length(component))) {
   for (j in component[[i]]) {
    if (length(lmajor[[j]]) >= min_connections) {
      component[[i]] <- append(component[[i]], lmajor[[j]])
      component[[i]] <- unique(component[[i]])
    }
  }

}

component <- unique(lapply(component, sort))
for (i in c(1:length(component))){
  for (j in c(1:length(component))) {
    if (component[[i]] %in% component[[j]]) {
      component[[i]] <- append(component[[i]], component[[j]])
      component[[i]] <- unique(component[[i]])
    }
  }
}
component
```

