---
title: "For Topology 4/10/2020"
author: "Fred Kaesmann"
output:
  pdf_document: default
  html_notebook: default
date: "4/7/2020"
---

From my notes from Monday (4/7/2020):

Example filter function:
\begin{equation*}
  \Sigma = \text{ Sample } = \{ ( x_i, y_i) \}
\end{equation*}


Program cover with resolution and gain 
\begin{itemize}
  \item If given a resolution and gain from user
  \item and some given height function (example above)
  \item Determine the overlap of covers (intersections of pullbacks)
  \item cluster points in overlap (user's choice)
  \item cluster become the connected components
\end{itemize}



Small program for Wednesday:
\begin{enumerate}
  \item Input:
  \begin{itemize}
    \item $N$: points in $R^2$ with $y \geq 0$
    \item $r$: resolution value
    \item $g$: gain
  \end{itemize}
  \item Output: Under $r$ and $g$ values given
  \begin{itemize}
    \item Produce the preimages of the covering intervals 
    \item "If you're feeling froggy:" Maybe consider an instersection matrix
            for when pullbacks nontrivially intersect?
  \end{itemize}
\end{enumerate}

\clearpage

## Part 1

```{r}
PullBackCovers <- function(x, y, resolution, gain){

  ##### Create New Data Frame from x and y Values Entered ####
  df <- data.frame(
    x = x,
    y = y
  )
  
  ##### Initial Info ####
  y_max <- max(df$y)                                 # When to stop
  increment <- resolution * (1 - gain)               # Lower bound iterations
  n_sets <- ceiling(y_max / increment)               # Number of Sets

  ##### Cover Boundaries ####
  # Initialize Variables
  lower <- 0 - increment
  upper <- lower + resolution
  covers <- list()
  for (cover_number in c(1:n_sets)) {
    covers[[cover_number]] <- df[df$y > lower & df$y < upper, ] # List entry cover
    lower <- (increment * cover_number)                         # Lower Bound
    upper <- lower + resolution                                 # Upper Bound
    
  }
  ##### Meaningful List Entry Names ####
  names(covers) <- paste0("Cover_", c(1:n_sets))
  
  ##### Return the list ####
  return(covers)
}


```



```{r}
covers <- PullBackCovers(runif(10)*10, runif(10)*10,
                           resolution =  4, 
                                 gain =  0.75)
print(covers[c(1:2)])
```


\clearpage

## Part 2
```{r}
CoverAdjacency <- function(covers, binary = TRUE) {
  ##### Adjacency Matrix ####

  ##### Packages ####
  library(dplyr, warn.conflicts = F, quietly = T)
  
  ##### Iteration Ranges ####
  list_length <- length(covers)
  list_range <- c(1:list_length)
  
  ##### Initialize Empty Matrix #####
  adj_mat <- matrix(nrow = list_length,
                   ncol = list_length,
                   dimnames = list(
                     paste0("C",list_range),
                     paste0("C",list_range)
                   ))
  
  ##### ith Row, jth Column ####
  for (i in list_range) {
  for (j in list_range) {
     
   ##### Binary Adjacency Matrix ####
   if (binary == TRUE) {
     if (nrow(anti_join(covers[[i]], covers[[j]], by = c("x", "y"))) < nrow(covers[[i]])){
       adj_mat[i, j] <- 1
     } else {
       adj_mat[i, j] <- 0
     }
    ##### Count the Overlapped Points ####
   } else {
     adj_mat[i, j] <- nrow(covers[[i]]) - nrow(anti_join(covers[[i]], covers[[j]], by = c("x", "y")))
   }
     
  }
  }
  
  ##### Return the Adjacency Matrix ####
  return(adj_mat)
}

```

\clearpage

```{r}
##### Binary Adjacency Matrix ####
CoverAdjacency(covers)
```

```{r}
##### Count the Points in Common ####
CoverAdjacency(covers, binary=FALSE)
```



