---
title: "Persistenc Diagram Construction Notebook"
output: html_notebook
---

\section{Persistence Diagram}
Map Births and Deaths of components in each cover as we move through "time."

We plot $(x_i, y_i)$ where $x_i$ indicates the cover in which we birth a new 
component, and $y_i$ idicates the death of a new component. 

\section{Pseudo Code}
| Increment by cover
|     A component born in the $ith$ cover 
|     and dies in the $jth$ cover 
|     gets a point at $(i, j)$ 


We begin with a vector of components 


```{r our_package, warning=FALSE, message=FALSE}
##### Packages ####
library(FredsVietorisRips)  # Our Very Own
library(magrittr)
library(dplyr)
library(ggplot2)
library(gridExtra)
library(reticulate)
```


```{r sourcepy}
##### Source our Python Code for Later Use ####
# All the code
filelist <- list.files("../python/")
for (f in filelist) {
  source_python(paste0("../python/", f))
}
```


```{r get_data}
##### Import Data ####
df_original <- read.csv("../data/Clark_Sample_data.csv", col.names = c("x", "y"))
```

```{r sample_data}
##### Take a section of the original data ####
sample_size <- 1000
df <- df_original[df_original$x < 0.5 & df_original$y < 0.25, ]
df <- sample_n(df, sample_size)

# Make it compatible with Pyton, start with zero
rownames(df) <- as.numeric(rownames(df)) - 1
```



```{r filter_by_cover}
##### Filter ####
resolution <- 0.04
gain <- 0.75
list_of_covers <- PullBackCovers(df$x, df$y, resolution, gain)
```


```{r dbscan}
##### Choose Epsilon and Minimum Connections ####
epsilon <- 0.1
min_connections <- 3

##### How many components do we have? #####
components <- vector(length = length(list_of_covers))
for (cover in c(1:length(list_of_covers))) {
  components[cover] <- Pairwisedist(list_of_covers[[cover]]$x, list_of_covers[[cover]]$y) %>%  # Create a distance matrix
    AdjacencyMatrix(epsilon) %>%                                                               # Determine Adjacent Points
    FredsDBSCAN(min_connections) %>%                                                           # Scan for Clusters
    length()                                                                                   # How many clusters?
}

components

components - lag(components)
```


























