---
title: "Two Simplices"
output: html_notebook
---

Given a set of points 
```{r}
df <- data.frame(
  x = runif(20),
  y = runif(20)
)
```

We start with a matrix of their pairwise distances
```{r}
pwdmat <- Pairwisedist(df$x, df$y)
```

From there we want a matrix of adjacencies. 
We determine adjacencies by which pairs of points are within a distance,
$\epsilon$.
These connections represent our set of 1-simplices.

For our purposes now, we want just the top right of the adjacency matrix or the
$\binom{N - 1}{2}$ pairs in the top right of the matrix, where $N$ is the total
number of point provided. 
```{r}
epsilon <- 0.5
adj_triangle <- AdjacencyMatrix(pwdmat, epsilon, top_right = TRUE)
```

To make things faster, we remove all points from our matrix who have less 
than 3 connections 

```{r}
adj_triples <- RemoveNoise(adj_triangle, 3)
```

Using the new matrix of points with connections $\geq 3$ we create a matrix
of three point triangles.

```{r}
triple_mat <- Connected_Triples(adj_triples)
```


With the new matrix of triangles, we create a newer matrix of triangles whose
points are all within $\epsilon / 2$ of the centroid of the triangle.

```{r}
two_simplices <- Two_Simplex(df$x, df$y, epsilon, triple_mat)
two_simplices
```










