---
title: "Wine Map"
author: "by Fred"
output:
  pdf_document: default
  html_notebook: default
---

```{r setup, include=FALSE}
##### Chunk Options ####
library(knitr)
opts_chunk$set(
  warning = F,
  message = F,
  fig.width = 6,
  fig.asp = 0.618
)
```

\section{What is the Mapper?}
\begin{enumerate}
  \item From data set $\mathbb{X}$
  \item We create a matrix of distances
  \item From a quick analysis of our distance matrix,
          we reasonably choose $\epsilon$
  \item By using our chosen $\epsilon$, we create an adjacency matrix
  \item We cover our space
  \begin{enumerate}
    \item We select a point at random
    \item We draw a ball of radius $\epsilon$ around it (initialize Cover $i$).
    \item Points within $\epsilon$ of our chosen point are allocated to Cover $i.$
    \item We select an uncovered/unallocated point and repeat the last two steps
            until all points have been covered.
  \end{enumerate}
  \item We create a new adjacency matrix of covers where two covers are adjacent 
          if their intersection is nonempty.
  \item We select additional information for coloring and sizing our nodes \\
        In this case:
          \begin{itemize}
            \item Coloring: Red Wines were given a value of 1, White Wines 0.
            Node Color was determined by mean value of wine type in each cover.
            \item Sizing: Nodes are sized by the number of points in each cover. 
          \end{itemize}
  \item Visualize it
\end{enumerate}


\section{Step 1: Packages and Data}

```{r packages}
##### Packages ####
library(tidyverse)    
library(igraph)
library(RColorBrewer)
library(FredsVietorisRips)
```

```{r data}
##### Import Data #### 
red <- read.csv("D:winequality-red.csv")
white <- read.csv("D:winequality-white.csv")

##### Add column Type ####
# Column of 
#   1's for Red
#   0's for White
red$type <- 1
white$type <- 0

##### Sample and Merge ####
set.seed(2020)
sample_size <- 50
df <- red %>%
  sample_n(sample_size) %>%
  full_join(sample_n(white, sample_size))
```

\section{Step 2: Distance Matrix}

```{r pairwise_distance}
##### Distance Matrix ####
# Euclidean distance, first 11 columns
# Not including the "Quality" indicator
d <- df[1:(ncol(df)-1)] %>%
  dist() %>%
  as.matrix()
```


\section{Step 3: Choose $\epsilon$}
This is where things get a bit subjective

We can look at a histogram of the distances between points to get a feel for
which $\epsilon$ might be most appropriate. 

```{r choose_epsilon}
##### Tidy Data ####
d_tidy <- TidyDistanceFrame(d)

##### Distance Distribution ####
d_tidy %>%
  ggplot(aes(Distance)) +
  geom_histogram(bins = 30) +
  ylab("Points within this Distance") +
  ggtitle("Distribution of Distances")
```

\subsection{"Epsilon Frame"}

We felt that just a histogram alone was inadequate to determine a sufficient
$\epsilon$, so instead we created a simple "Epsilon Frame," where we track
the number of connections made out of the total possible for a given set,
$\binom{n}{2},$ as well as the number of components present.

```{r epsilon_frame}
##### Create our "Epsilon Frame" ####
# From a vector of epsilon values
# we determine 
# 1. The proportion of connections made 
# 2. The number of components present
eframe <- CreateEpsilonFrame(d, seq(0, 50, by = 5))

##### Print Fancy Table ####
eframe %>% 
  kable(caption = "Epsilon Frame")
```


```{r epsilon_curves}
##### Visualize Curves ####
eframe %>%
  mutate(Components = Components / 100) %>%
  ggplot(aes(x = Epsilon)) +
  geom_line(aes(y = Connections), color = "red") + 
  geom_line(aes(y = Components), color = "blue") +
  ylab(NULL) +
  ggtitle("Epsilon Curves",
          "Connections in Red, Components in Blue")
```




\section{Step 2: Open Covers}
From our Epsilon Frame above, we opted to go with 25

```{r open_covers}
epsilon <- 25
  ##### Open Covers ####
  # Distance matrix -> Adjacency Matrix -> Open Covers
  covers <- d %>%
    AdjacencyMatrix(epsilon) %>%
    OpenCoverEballs()
  
  ##### Adjacent Covers ####
  # Create a graph from an 
  # Adjacency matrix of our covers
  covers.plot <- covers %>%
    CoverAdjacencies() %>%
    graph_from_adjacency_matrix(mode="undirected")


```



\section{Step 3: Plot it}


```{r plot}
  ##### Graph ####
    ##### Node Features ####
    # For each of our Covers (which becomes each of our Nodes)
    #   1) Turn the average number of Red wines in each Cover
    #       into an integer [0-11] + 1 to color the Node
    #      a) A Node of White wine only = 1 (white)
    #      b) A Node of Red wine only = 12 (darkpurple)
    #      c) All other Nodes are colored on a scale of [2-11]
    #   2) Size the each Node by the number of points (wines)
    #       in each Cover
    covers.size <- vector()
    covers.color <- vector()
    
    for (i in 1:length(covers)) {
        # No Red Wines
      if ( mean( df$type[ covers[[i]] ] ) == 0 ) {
        covers.color[i] <- 1
        # No White Wines
      } else if ( mean (df$type[ covers[[i]] ] ) == 1) {
        covers.color[i] <- 12
        # All other Nodes
      } else {
        covers.color[i] <- round( mean( df$type[ covers[[i]] ]) * 10 ) + 1
      }
      
      covers.size[i] <- length(covers[[i]])
    }
    
    ##### Coloring ####
    redscale <- c("white", brewer.pal(9, "Reds"), "violetred3", "violetred4")
    V(covers.plot)$color <- redscale[covers.color]
    
    ##### Sizing ####
    V(covers.plot)$size <- log(covers.size) * 10
    
    ##### Plot ####
    plot(covers.plot)
```



\section{Rerun}

```{r rerun_function, echo=FALSE}
RerunMapper <- function(sample_size = 50,
                            epsilon = 25,
                              title = NULL) {

  ##### Data ####
  
    ##### Sample and Merge ####
    df <- red %>%
      sample_n(sample_size) %>%
      full_join(sample_n(white, sample_size))
    
  ##### Topology #### 
    ##### Distance Matrix ####
    # Euclidean distance, first 11 columns
    # Not including the "Quality" indicator
    d <- df[1:(ncol(df)-1)] %>%
      dist() %>%
      as.matrix()
    
    ##### Open Covers ####
    # Distance matrix -> Adjacency Matrix -> Open Covers
    covers <- d %>%
      AdjacencyMatrix(epsilon) %>%
      OpenCoverEballs()
    
    ##### Adjacent Covers ####
    # Create a graph from an 
    # Adjacency matrix of our covers
    covers.plot <- covers %>%
      CoverAdjacencies() %>%
      graph_from_adjacency_matrix(mode="undirected")
  
  ##### Graph ####
    ##### Node Features ####
    # For each of our Covers (which becomes each of our Nodes)
    #   1) Turn the average number of Red wines in each Cover
    #       into an integer [0-11] + 1 to color the Node
    #      a) A Node of White wine only = 1 
    #      b) A Node of Red wine only = 12 
    #      c) All other Nodes are colored on a scale of [2-11]
    #   2) Size the each Node by the number of points (wines)
    #       in each Cover
    covers.size <- vector()
    covers.color <- vector()
    
    for (i in 1:length(covers)) {
        # No Red Wines
      if ( mean( df$type[ covers[[i]] ] ) == 0 ) {
        covers.color[i] <- 1
        # No White Wines
      } else if ( mean (df$type[ covers[[i]] ] ) == 1) {
        covers.color[i] <- 12
        # All other Nodes
      } else {
        covers.color[i] <- round( mean( df$type[ covers[[i]] ]) * 10 ) + 1
      }
      
      covers.size[i] <- length(covers[[i]])
    }
    
    ##### Coloring ####
    redscale <- c("white", brewer.pal(9, "Reds"), "violetred3", "violetred4")
    V(covers.plot)$color <- redscale[covers.color]
    
    ##### Sizing ####
    V(covers.plot)$size <- log(covers.size) * 10
    
    ##### Plot ####
    plot(covers.plot, main = title)
}
```

```{r rerun}
##### Rerun the Mapper ####
# Each iteration takes a new sample of both data sets
map_iterations <- 10
for (i in 1:map_iterations) RerunMapper(sample_size = 100,
                                        title = paste("Map", i))

```



