---
title: "README"
subtitle: 'A "How To"'
author: "Fred Kaesmann, Dani Rosenberg and Kun Woo Lee"
date: "April 4, 2020"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,
                      message = FALSE,
                      warning = FALSE,
                      # Figure Settings
                      fig.asp = 0.618,
                      fig.width = 6)
```

## How to Handle our Package
Welcome to the intro to
```{r}

library(FredsVietorisRips)
```
We lovingly designed this package for use in Dr. Aaron Clark's MAT 499
Topological Data Analysis Independent Study.
\footnote{TDAIS for short}
This package was developed to replicate aspects of the Mapper Algorithm from 
\textit{insert sing, carlsson et al}.

It is to be used in the following manner

## Usage

### Downloading
To download from Github, copy and paste the following into the console
```{r eval=FALSE}
if (!require(FredsVietorisRips)) install_github("ftkjr/FredsVietorisRips")
library(FredsVietorisRips)
```

### Simplices, 0 and 1 

To create our 1-simplices,
\begin{enumerate}
  \item[Step 1:] We generate a data frame of random $x$ and $y$ 
coordinates 

  \item[Step 2:] We create a matrix of pairwise distances. 
                  Each cell $i,j$ represents the distance from the $i$th point
                  to the $j$th point.
  
  \item[Step 3:] Given a distance, $\epsilon,$ we return an 
                  adjacency matrix which displays a 1 for each pair whose 
                  Euclidean Distance is within the given $\epsilon$ value
  
  \item[Step 4:] From the adjacency matrix, we pull out the points which are 
                  adjacent.
                  This is a list of all 1-simplexes in the dataset
                  
  \item[Step 5:] Last, we visualize the simplices using the `ggplot2` library
\end{enumerate}


```{r step_1}
##### Step 1: Create Data Frame ####
frame_size <- 100
df <- data.frame(
  x = runif(frame_size),
  y = runif(frame_size),
  Point = paste0("P", c(1:frame_size))
)

##### Step 2: Pairwise Distance Matrix ####

pwdmat <- Pairwisedist(df$x, df$y)
# pwdmat

##### Step 3: Given epsilon, determine Adjacency ####
# Given epsilon
epsilon <- 0.5

# Determine Adjacency 
adjacency_matrix <- AdjacencyMatrix(pwdmat, epsilon)
# adjacency_matrix


##### Step 4: Which Points are Adjacent? ####
paired_points <- AdjacentPairs(adjacency_matrix)
# paired_points

##### Step 5: Plot 0 and 1 Simplices ####
Plot_1_Simplices(df$x, df$y, epsilon)


```











